animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation

### 2025-07-08 response_to_review.qmd

- *When using custom breaks, the error displays "The breaks vector must have the same number of groups as ncat argument", but it is stated in the article that the length of breaks need to equal the number of categories plus 1.*

Response: The code has been fixed.

- *When running the code on an M1 MacBook Pro, some of the shaded sections tear apart and look like white circles. This happens particularly with the wallaby plots.*

Response: ??? It might be related to the default colour for shaded sections. Will need examples to replicate.

@@

- *Are there plans to include additional visuals to help keep track of the points as they disappear? In the NYT article, the points are summed and proportions are calculated as the points reach the right side of the plot and I think this would be a nice addition in the future.*

Response: ??? Not at the moment since it can lead to a visual overload.

- *When there are many data points, it becomes difficult track the general movement of observations. Would there be any options to trace the paths to show where the points have moved (like a line with low alpha)?*

Response: The gganimate elements (e.g., shadow trail) work normally with this package. We have added an explanation on page 8.

@@

Response: Fixed

### *Code*

*I did not find the commands in the package very user-friendly. For instance, to produce a plot I must go through many functions, prior to typing the plot to appear. The example in the aeles dataset is such a case. I need to type 4 commands in order to see a plot, which I repeat, cannot understand, nor interpret easily. I would prefer a function that does all this preprocessing and plots the data. Legends are missing, and in my opinion, the help files are not very helpful for someone who wants to use this package easily, to visualize their data.*

Response: ??? It was a design choice to allow for flexibility. For example, to add ggplot or gganimate elements, it needs to be done between the processes. There is also a page for getting started, which discusses it in more detail. (https://krisanata.github.io/animbook/articles/animbook.html)[https://krisanata.github.io/animbook/articles/animbook.html]

### 2025-07-08 animbook-journal.Rmd
first review draft

---
title: "animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation"
abstract: >
  Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys," we provide new visualization functions to generate animations that display this type of data in an R package called animbook. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_pdf_article:
    toc: no
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
bibliography: animbook-journal.bib
---

@@

  The political data that exhibits the movement of voters switching party affiliations between elections can be a valuable insight into the behavior of the voters. This data could be extended to incorporate demographic information about the voters, providing analysts with significant insight into voter behavior and allowing them to consolidate effective campaigns for their political party. It can also be applied to marketing data, where customers shift their product interest to the competitor, providing the marketing analysts with an understanding of both the company's products and the overall market.

@@

It is necessary to store the plot as a `ggplot` object before passing it to the final function, `anim_animate()`, to animate. This function will automatically detect which rendering method was specified in the previous steps and add the minimum requirement functions accordingly. By default, if the user specifies the rendering as `gganimate`, then it will add the `transition_time()` function from the \CRANpkg{gganimate} package. Otherwise, the `animation_opts()` function will be added from the \CRANpkg{plotly} package.

@@

I would like to acknowledge Chika Saka for providing the inspiration that led to the development of this R package. Additionally, thanks to Masayuki Jimichi for the efforts in collecting the sample Osiris data, which contributed to the practical application of the package.

>>>>

---
title: "animbook: Visualizing Changes in Performance Measures and Demographic Affiliations using Animation"
abstract: >
  Performance is often reported in quantiles or ranks and often repeatedly provided over time. Between elections, voters often switch party preferences, which produces categorical values over time. Inspired by the animation in the New York Times article "Extensive data show punishing reach of racism for black boys," we provide new visualization functions to generate animations that display this type of data in an R package called animbook. Functions to pre-process the data and prepare it for the animation are provided. The methods will be suitable generally for data where categorical variables are measured over time, and provide a way to engagingly view changes.
author:  
  - name: Krisanat Anukarnsakulchularp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    - Melbourne, Australia
    orcid: 0009-0008-5638-7124
    email:  kanu0003@student.monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: 
      - Department of Econometrics and Business Statistics
      - Melbourne, Australia
    email: dicook@monash.edu
    orcid: 0000-0002-3813-7155
type: package
output: 
  rjtools::rjournal_web_article:
    self_contained: yes
    toc: no
  rjtools::rjournal_pdf_article:
    toc: no
bibliography: animbook-journal.bib
---

@@

  The political data that exhibits the movement of voters switching party affiliations between elections can be a valuable insight into the behavior of the voters. These data could be extended to incorporate demographic information about the voters, providing analysts with significant insight into voter behavior and allowing them to consolidate effective campaigns for their political party. It can also be applied to marketing data, where customers shift their product interest to the competitor, providing the marketing analysts with an understanding of both the company's products and the overall market.

@@

It is necessary to store the plot as a `ggplot` object before passing it to the final function, `anim_animate()`, to animate. This function will automatically detect which rendering method was specified in the previous steps and add the minimum requirement functions accordingly. By default, if the user specifies the rendering as `gganimate`, then it will add the `transition_time()` function from the \CRANpkg{gganimate} package. It also means that the functionality from the \CRANpkg{gganimate} package (e.g., `shadow_trail()` function) will work by default. Otherwise, the `animation_opts()` function will be added from the \CRANpkg{plotly} package.

@@

We would like to acknowledge Chika Saka for providing the inspiration that led to the development of this R package. Additionally, thanks to Masayuki Jimichi for the efforts in collecting the sample Osiris data, which contributed to the practical application of the package.
